---
title: "Vaccine effectiveness report"
output: 
  html_document:
    code_folding: hide
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library('tidyverse')
library('lubridate')
library('gt')
library('patchwork')

source(here::here("lib", "html_output_functions.R"))

output_dir <- here::here("released_output")

summary_stats <- read_rds(file.path(output_dir, "data_summary", "summary_stats.rds"))

run_date <- "2021-01-20" ## need to update this once metadata is available to access
end_date <- summary_stats$end_date

pct <- scales::label_percent(accuracy=0.1)


bar_theme <- 
  theme_bw(base_size = 9) + 
  theme(
    panel.border = element_blank(), #axis.line.x = element_line(colour = "black"),
    panel.grid = element_blank(),
    panel.grid.major.x = element_line(colour="grey"),
    axis.ticks = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0),
    plot.title = element_text(hjust = 0),
    plot.title.position = "plot",
    plot.caption.position =  "plot",
    plot.caption = element_text(hjust = 0, face= "italic")
  )



```

## Vaccine effectiveness in OpenSAFELY-TPP data

This report examines covid-related outcomes in TPP-registered patients who have received a vaccine for SARS-CoV-2. 

Outcomes include:

* Positive Covid case identification in primary care, using the following clinical codes:
  * [probable covid](https://codelists.opensafely.org/codelist/opensafely/covid-identification-in-primary-care-probable-covid-clinical-code/2020-07-16)
  * [positive covid test](https://codelists.opensafely.org/codelist/opensafely/covid-identification-in-primary-care-probable-covid-positive-test/2020-07-16/)
  * [covid sequelae](https://codelists.opensafely.org/codelist/opensafely/covid-identification-in-primary-care-probable-covid-sequelae/2020-07-16/)
* Positive SARS-CoV-2 test, as reported via the Second Generation Surveillance System (SGSS)
* Covid-related hospital admission, where [Covid](https://codelists.opensafely.org/codelist/opensafely/covid-identification/2020-06-03/) is listed as a reason for admission.
* Covid-related death, where [Covid](https://codelists.opensafely.org/codelist/opensafely/covid-identification/2020-06-03/) is mentioned anywhere on the death certificate.
* All-cause death.

### Data notes

The dataset used for this report was created using the study definition `/analysis/study_definition_delivery.py`, using codelists referenced in `/codelists/codelists.txt`.

The dataset is based on OpenSAFELY-TPP data extracted on `r run_date`, with event dates occurring no later than `r end_date`.

```{r tte_rate_function, echo=FALSE}

print_event_rates <- function(variable,  variable_name){

  table <- read_csv(
    file.path(output_dir, "tte", "tables", paste0("event_rates_at_7days_", variable, ".csv")),
    col_types = "fiididididid",
    na="-"
  )
  names(table)[1] <- variable_name
  
  renamer <-
  c(
    posPC_ = "primary care case_", 
    posSGSS_ = "positive test_",
    admitted_ = "covid-related admission_",
    coviddeath_ = "covid-related death_",
    `death_n` = "all-cause death_n",
    `death_%` = "all-cause death_%"
  )
  
  table <- rename_with(table, ~str_replace_all(string = .x, renamer))
  
  gt_table <- table %>%
  gt() %>%
  tab_spanner_delim(delim="_", gather=TRUE) %>%
    # gather doesn't work!
    fmt_percent(
      columns = ends_with(c("%")),
      decimals = 1
    ) %>%
    fmt_missing(
      everything(),
      missing_text="-"
    ) %>%
    tab_header(
      title = md("Post-vaccination event rates at 7-days")
    ) %>%
    tab_source_note(
      "- indicates redacted values"
    )

  gt_table
}


```


```{r vaccinetype_function, echo=FALSE}

print_vaccinetype <- function(variable,  variable_name){

  table <- read_csv(
    file.path(output_dir, "vaccine_type", "tables", paste0("vaccine_type_", variable, ".csv")),
    col_types = "ffiddl",
    na="-"
  )
  
  gt_catcat(summary_catcat=table, 
            var1_name = variable_name, 
            var2_name = "vaccine type", 
            title="Vaccine type", 
            source_note = "- indicates redacted values",
            pct_decimal=1
            )
}
```

```{r tte_barplot_function, echo=FALSE}

print_variable_freq <- function(variable, variable_name){
  
  table <- read_csv(
    file.path(output_dir, "variable_summary", "tables", paste0("categorical_", variable, ".csv")),
    col_types = cols(
      n=col_integer(),
      pct=col_double(),
      pct_nonmiss=col_double()
    )
  )
  
  ywrapwidth <- 50
  
  plotdata <- table %>%
    mutate(
      variable_explicit_na=factor(str_wrap(.level, ywrapwidth), levels=str_wrap(unique(.level), ywrapwidth))
    )
  
  plot <- ggplot(plotdata)+
    geom_bar(aes(x=n, y=variable_explicit_na), stat='identity', width=0.8)+
    geom_text(aes(x=n, y=variable_explicit_na, label=pct(pct)), hjust=-0.2)+
    #scale_x_continuous(limits=c(0, xlim_upper), breaks=seq(0,10000,1000))+
    scale_y_discrete(limits = rev(levels(plotdata$variable_explicit_na)))+
    labs(
      #title=variable_name,
      #subtitle = subtitle,
      x="Patients", 
      y=NULL
    )+
    bar_theme+
    NULL
  
  print(plot)
  
}

```


```{r tte_survplot_function, echo=FALSE}
  #cat("  ")
  
print_variable_surv <- function(variable, variable_name){
  
  knitr::include_graphics(
    here::here(output_dir, "tte", "figures", paste0("plot_patch_",variable, ".svg"))
  )
}

```


## Summary

As of `r end_date`, there were `r summary_stats$total_vaccinated` TPP-registered patients who had received at least one dose of a vaccine for covid. Of these, 
`r summary_stats$total_vaccinated_pfizer` received the Pfizer vaccine and `r summary_stats$total_vaccinated_oxford` received the Oxford/AZ vaccine.

## Summaries by patient characteristics

The tables below show, for each characteristic:

* Vaccine type, including those who have received both vaccine types
* 7-day post-vaccination event rates
* Cumulative post-vaccination event rates over time.

### Age

```{r age, echo=FALSE, out.width='90%', results='asis'}
print_vaccinetype("ageband", "Age")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_event_rates("ageband", "Age")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_variable_surv("ageband", "Age")
cat("  \n", "  \n", "  \n")
```


### Sex

```{r sex, echo=FALSE, out.width='90%', results='asis'}
print_vaccinetype("sex", "Sex")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_event_rates("sex", "Sex")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_variable_surv("sex", "Sex")
cat("  \n", "  \n", "  \n")
```

### Ethnicity

```{r ethnicity, echo=FALSE, out.width='90%', results='asis'}
print_vaccinetype("ethnicity", "Ethnicity")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_event_rates("ethnicity", "Ethnicity")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_variable_surv("ethnicity", "Ethnicity")
cat("  \n", "  \n", "  \n")
```

### Index of Multiple Deprivation (IMD)

```{r imd, echo=FALSE, out.width='90%', results='asis'}
print_vaccinetype("imd", "IMD")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_event_rates("imd", "IMD")
cat("  \n", "  \n", "  \n", "   \n")
cat("##  \n")
print_variable_surv("imd", "IMD")
cat("  \n", "  \n", "  \n")
```

### STP

```{r stp, echo=FALSE, out.width='90%', results='asis'}
print_vaccinetype("stp", "STP")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_event_rates("stp", "STP")
cat("  \n", "  \n", "  \n")
cat("##  \n")
print_variable_surv("stp", "STP")
cat("  \n", "  \n", "  \n")
```
